\section{Introduction}
This chapter describes the implementation of RV32ICore in VHDL, covering the structural organization of the design, the implementation of each major component, key design decisions made during development, and the toolchain used to compile and load test programs into the simulation environment. The implementation follows directly from the architecture described in Chapter 3, with each architectural unit realized as a distinct VHDL entity.    
\section{VHDL Design Organization}
RV32ICore is organized as a hierarchy of VHDL entities, each encapsulating a well-defined functional unit. The top-level entity \textit{P\_Processor\_Hazard} instantiates three major components: the datapath \textit{(P\_Datapath)}, the control unit \textit{(P\_Control\_Unit)}, and the hazard unit \textit{(Hazard\_Unit)}. Memory is intentionally kept external to the processor entity, treated as a peripheral connected through the top-level ports. This separation reflects a deliberate design decision: by exposing the instruction and data memory interfaces at the top level, the processor can be connected to different memory configurations without modifying its internal logic, improving modularity and testability.

The top-level interface exposes the following ports: a clock and reset, a 32-bit instruction input, a 32-bit read data input from data memory, and outputs for the program counter, data address, write data, memory write enable, byte enable, and access mode signals. This interface is sufficient to connect RV32ICore to any external Harvard memory system.
\section{Pipeline Register Implementation}
A central implementation decision in RV32ICore is the use of a single generic VHDL component, \textit{RegEn}, to implement all pipeline registers throughout the design. This component is parameterized by data width through a generic and supports four operating modes controlled by its input signals: normal clocked operation, stall (output held, input ignored), flush (output loaded with a predefined NOP value), and synchronous reset (output cleared to zero).

The NOP value inserted during a flush is x"\textit{00000033}", corresponding to the RV32I instruction \textit{ADD x0, x0, x0}. This is a valid instruction that writes to register x0, which is hardwired to zero, making it functionally harmless. Using a valid instruction rather than an all-zero pattern avoids potential undefined behavior in the decode and control logic during pipeline recovery.

Separate width variants of the same component (\textit{RegEn\_5bit, RegEn\_9bit, RegEn\_6bit}) are used for narrower signals such as register addresses and reduced control buses, maintaining the same structural pattern throughout the design.
\section{Control Unit Implementation}
The control unit is implemented as \textit{P\_Control\_Unit} and is itself structured as a hierarchy of sub-components. It contains three functional blocks: the main control unit (\textit{P\_Main\_unit}), the ALU control unit (\textit{Alu\_cntr}), and the execution flow controller (\textit{Excution\_Flow}).

\textit{P\_Main\_unit} is a combinational decoder that maps the opcode and Funct3 fields to the full set of control signals required by the datapath. Its outputs include the immediate extension select, memory access width, register write enable, memory write enable, memory extension control, write-back select, PC select, ALU input select, branch mode, and ALU input override for AUIPC and LUI instructions.

\textit{Alu\_cntr} decodes the opcode, Funct3, and bit 5 of Funct7 to produce the 4-bit ALU operation code. Using only bit 5 of Funct7 is sufficient to distinguish between arithmetic pairs such as ADD/SUB and SRL/SRA, avoiding the need to decode all seven Funct7 bits.

\textit{Excution\_Flow} resolves the final PC select signal in the Execute stage. It receives the initial PC select from the control bus pipeline register, the branch mode encoding, the ALU zero flag, and the ALU LSB, and produces the final 2-bit PC select that determines whether the next PC is PC+4, the ALU result (for JALR), or the branch target address.
\subsection{Control Bus Pipelining}
A notable implementation strategy in the control unit is the aggregation of all control signals into a single control bus that is pipelined through\textit{RegEn} registers in parallel with the datapath pipeline registers. Rather than routing individual control signals through separate pipeline registers, the control signals are packed into a 21-bit bus at the Decode stage output and propagated as a unit through successive pipeline stages. At each stage boundary, only the subset of control signals relevant to that stage is extracted from the bus.

The bus is progressively narrowed as signals are consumed: the full 21-bit bus is reduced to a 9-bit bus entering the Memory stage (retaining only memory and write-back controls), and further reduced to a 6-bit bus entering the Write-Back stage (retaining only register write enable, memory extension control, and write-back select). This approach simplifies the pipeline register instantiation and makes the control signal flow explicit and traceable in the VHDL source.
\section{Datapath Implementation}
The datapath is implemented as \textit{P\_Datapath} and instantiates all pipeline stages, pipeline registers, forwarding multiplexers, and functional units. The following describes the implementation of each stage.
\subsection{Fetch Stage}
The PC is stored in a \textit{RegEn} instance configured with the stall input connected to \textit{StallF}. Under normal operation it loads \textit{Pc\_Next} on every rising clock edge. The \textit{Pc\_Next} signal is selected by a three-input multiplexer (\textit{Mux3}) among three sources: the ALU result (for JALR), PC+4 (for sequential execution), and the branch target address computed in the Execute stage (for taken branches). PC+4 is computed by a dedicated adder that adds the constant four to the current PC. Both the current PC and PC+4 are registered into the IF/ID pipeline register for use in later stages.
\subsection{Decode Stage}
The register file is read combinationally using rs1 and rs2 addresses extracted directly from the pipelined instruction word at bits [19:15] and [24:20] respectively. The destination register address rd is extracted from bits [11:7] and propagated through the pipeline in a dedicated 5-bit pipeline register chain (\textit{Reg\_A3\_PLR3} through \textit{Reg\_A3\_PLR5}) alongside the instruction, reaching the Write-Back stage where it addresses the register file write port.

The immediate extension unit receives the full 32-bit instruction word and a 3-bit select signal from the control unit, and produces the appropriate 32-bit sign-extended immediate for the instruction's format.

The register source addresses rs1 and rs2 are also registered separately into Rs1E and Rs2E for use by the Hazard Unit in forwarding comparisons.
\subsection{Execute Stage}
Two forwarding multiplexers (\textit{ForwardingA} and \textit{ForwardingB}) sit at the inputs of the ALU input selection multiplexers. Each is a Mux3 instance that selects among the register file output (no forwarding), the current Write-Back result (forwarding from WB), and the EX/MEM ALU result (forwarding from MEM). The selection is controlled by the 2-bit \textit{ForwardA} and \textit{ForwardB} signals from the Hazard Unit.

The forwarded rs2 value (\textit{ForwarededB}) is also registered into the MEM stage in a dedicated pipeline register (\textit{ForwaredB\_PLR4}). This is necessary for Store instructions, which require the value of rs2 at the Memory stage as the data to be written to memory, while the ALU computes the store address using rs1 and the immediate.

ALU input A is selected by \textit{ALU\_in\_A} (\textit{Mux3}) among the forwarded rs1 value, the PC (for AUIPC), and zero (for LUI). ALU input B is selected by \textit{ALU\_in\_B }(\textit{Mux2}) between the sign-extended immediate and the forwarded rs2 value.

The branch target address is computed by a dedicated adder summing the PC value carried from the Decode stage with the sign-extended immediate.
\subsection{Memory Stage}
The data address presented to external memory is the ALU result from the EX/MEM pipeline register. The write data is the forwarded rs2 value from \textit{ForwaredB\_PLR4}. The two least significant bits of the ALU result (\textit{Alu\_LSB\_Mem}) are passed to the top level as the ByteEn signal, providing the byte offset within a word to the external memory for sub-word accesses.
\subsection{Write-Back Stage}
The Memory Extension unit (\textit{Mem\_SignExt}) receives the raw 32-bit read data from memory, the two LSBs of the ALU result as a byte offset, and the memory extension control signal derived from Funct3. It produces a correctly sign- or zero-extended 32-bit value for byte and halfword load instructions.

The Write-Back multiplexer (\textit{Mux3}) selects among PC+4 (for JAL/JALR return address), the extended memory data (for load instructions), and the ALU result (for all other instructions). The selected value is written to the register file at the address held in \textit{Reg\_A3\_PLR5}.
\section{Hazard Unit Implementation}
The Hazard Unit is implemented as a fully combinational entity Hazard\_Unit. It receives the source register addresses from the Decode and Execute stages, the destination register addresses from the Execute, Memory, and Write-Back stages, the register write enables from the Memory and Write-Back stages, and the write-back control LSB and PC select MSB from the control pipeline.
\subsection{Forwarding Logic}
Forwarding priority is implemented as a priority-encoded conditional check. For each ALU input, the MEM stage forwarding path takes priority over the WB stage path, ensuring that the most recently computed value is always selected in the case of back-to-back dependent instructions. Forwarding is suppressed when the destination register is x0, since x0 is hardwired to zero and forwarding from it would produce incorrect results.

The forwarding condition also checks a combined signal: forwarding is enabled when either \textit{RegWriteM} is asserted or when the LSB of \textit{Pc\_Select\_Initial} is '\textit{0}', which identifies JAL and JALR instructions that write PC+4 to a destination register but are not flagged by the standard \textit{RegWrite} path at that pipeline stage.
\subsection{Load-Use Stall}
The load-use stall is detected by checking whether the \textit{Writeback\_cntr} LSB is asserted — indicating a load instruction in the Execute stage — and whether the load destination register matches either source register of the instruction currently in the Decode stage. When detected, \textit{StallF} and \textit{StallD} are asserted to freeze the Fetch and Decode stage pipeline registers, and \textit{FlushE} is asserted to insert a NOP into the Execute stage for one cycle.
\subsection{Control Hazard Flush}
The flush signal for taken branches is derived from the MSB of the \textit{Pc\_Select} signal, which is asserted whenever the processor takes a branch or jump. \textit{FlushE} is asserted as the logical OR of the load-use stall and the PC select signal, flushing the Execute stage in both cases. \textit{FlushD} is asserted on the PC select signal alone, flushing the Decode stage on taken branches. This correctly discards the two instructions fetched after the branch while leaving the Fetch stage free to begin fetching from the branch target on the following cycle.
\section{Toolchain and Program Loading}
To verify RV32ICore through simulation, a software toolchain is used to compile C test programs into binary executables targeting the RV32I ISA, which are then converted into a format that can be initialized into the instruction memory for simulation. The following code will be used as an Example:
\begin{minted}{C}
   // test.c

// Tests loops and addition
int sum_array(int arr[], int len) {
    int sum = 0;
    for (int i = 0; i < len; i++) {
        sum += arr[i];
    }
    return sum;
}
int main() {
    int arr[] = {3, 7, 2, 9, 1};

    int s = sum_array(arr, 5);       // expected: 22

    return 0;
}
\end{minted}

The compiler used is the xPack RISC-V Embedded GCC toolchain (\textit{riscv-none-elf-gcc}), configured to target the RV32I base ISA without compressed instructions or floating-point extensions, matching exactly the instruction set implemented by RV32ICore. A custom linker script defines the memory layout, setting the origin of the text segment to address \textit{0x00000000} to match the reset vector of the processor, and configuring the initial stack pointer to the top of the available data memory region.

A startup assembly file initializes the processor state before transferring control to the C main function. It sets the stack pointer register and performs any necessary early initialization before the C runtime environment is available.
\begin{minted}{riscv}
    00000000 <_start>:
   0:   40000113                li      sp,1024
   4:   11400513                li      a0,276
   8:   11400593                li      a1,276

0000000c <bss_loop>:
   c:   00b55863                bge     a0,a1,1c <bss_done>
  10:   00052023                sw      zero,0(a0)
  14:   00450513                addi    a0,a0,4
  18:   ff5ff06f                j       c <bss_loop>

0000001c <bss_done>:
  1c:   07c000ef                jal     98 <main>

00000020 <halt>:
  20:   0000006f                j       20 <halt>
\end{minted}

\textbf{Note:} The compiled code is very large hence a sample is given, and the remaining is illustrated in the Appendix.

