\paragraph{}
This project has presented the design, implementation, and verification of RV32ICore, a pipelined 32-bit processor implementing the RV32I Base Integer ISA, described in VHDL and verified through simulation. 
\paragraph{}
The work began with a study of the RISC-V ISA and foundational processor design concepts, establishing the academic context within which the project is situated. The architectural design of RV32ICore was then developed, comprising a classical five-stage Harvard pipeline with a dedicated hazard unit responsible for resolving data hazards through forwarding and stalling, and control hazards through pipeline flushing. The implementation translated this architecture directly into a hierarchy of VHDL entities, with notable design decisions including a generic reusable pipeline register component, a progressively narrowed control bus, and an external memory interface that improves modularity. The processor was verified by compiling and executing C programs of increasing complexity in ModelSim, with results confirmed against expected values. The full RV32I Base Integer ISA was demonstrated to execute correctly under simulation.
\paragraph{}
The project objectives stated in Chapter 1 have been met: a complete RV32I processor has been implemented in VHDL, verified through simulation using compiler-generated test programs, and documented to a standard suitable for academic presentation.
\paragraph{}
Several directions for future work are identified. The most immediate extension is deployment on an FPGA target, which would allow measurement of actual clock frequency, resource utilization, and real-world performance. Adding the RV32M extension would introduce hardware multiply and divide instructions, significantly broadening the range of programs the processor can efficiently execute. Implementation of the RISC-V privileged architecture specification — specifically Supervisor mode and the virtual memory system — would enable a full operating system to run on the processor, representing a substantial step toward a complete soft-core system. Finally, pipeline optimization through branch prediction or deeper pipelining could improve instruction throughput beyond what the current flush-on-branch approach achieves.