%\begin{}
\paragraph{ }
Network traffic volumes and attack sophistication have grown to a point where software intrusion detection systems face fundamental throughput limitations. A software IDS running on a general-purpose CPU must copy packet data through the OS network stack, schedule the inspection process, and share the CPU with other tasks---all of which introduce variable latency and reduce the fraction of packets that can actually be inspected at high line rates.
\paragraph{}
Reconfigurable hardware offers a different trade-off: inspection logic is implemented directly as digital circuits in an FPGA, so packets are examined as bytes arrive, with no operating system, no scheduling overhead, and no data copies. The cost is development complexity; the benefit is deterministic, wire-speed inspection.
\paragraph{}
This project builds a hardware IDS in VHDL, targeting the MicroPhase Z7-Lite development board (Zynq-7020, XC7Z020-1CLG400C), and integrates it with a RISC-V processor built independently by a collaborator. The Zynq device is a natural fit: its ARM Cortex-A9 Processing System (PS) receives Ethernet packets through the on-chip GigE MAC and can forward raw frame data to the Programmable Logic (PL) fabric via AXI-Stream DMA, where the IDS pipeline runs. The RISC-V processor also resides in the PL fabric and communicates with the IDS through a shared memory-mapped register bank.
\paragraph{}
The IDS is developed and validated entirely through behavioural simulation before any hardware integration, allowing each layer of the protocol stack to be verified independently with injected test packets.
