\documentclass[16pt]{mythesis}
\usepackage{booktabs}
\usepackage{float}
\usepackage{tikz}
\usepackage{tabto} 
\usepackage{minted}
\setminted{
    linenos=true,     % Add line numbers
    breaklines=true,  % Wrap long lines
    frame=lines,      % Add lines above and below the code
    fontsize=\footnotesize
}
\def\checkmark{\tikz\fill[scale=0.4](0,.35) -- (.25,0) -- (1,.7) -- (.25,.15) -- cycle;}  


\newcommand{\mynote}[3]{
    \fbox{\bfseries\sffamily\scriptsize#1}
    {\small$\blacktriangleright$\textsf{\emph{\color{#3}{#2}}}$\blacktriangleleft$}}
\newcommand{\yb}[1]{\mynote{Yahia}{#1}{red}}
%\usepackage{arabtex}

\input{preamble/packages}
\input{preamble/bib-setup}
\input{preamble/details}


\begin{document}
\maketitle

% Input anything that can go before the acknowledgements 

%\input{chapters/blank-pages.tex}

% Then the acknowledgements
\input{chapters/dedication}

\input{chapters/acknowledgement}




\input{chapters/abstractEn}


%\input{chapters/abstractAr}
% Print out the table of tables and table of figures and
% tell the template we're about to start the body of the
% thesis.

\thesisTables


\input{chapters/Abbreviations}
%\thesisBodyStart

% start of thesis body
% ---------------------------
%\Include {introduction}






% Print out the table of tables and table of figures and
% tell the template we're about to start the body of the
% thesis.
%\thesisTab  
\thesisBodyStart

% start of thesis body
% ---------------------------
%\Include {introduction}

\chapter{Introduction}
\input{chapters/introduction}

% Literature review / Background / Related work
\chapter{RISC-V Theoretical Background \& System Architecture}
\input{chapters/Chapter2}


\chapter{Implementation \& Verification}
\input{chapters/Chapter3}


\chapter{IDS Theoretical Background \& System Architecture}
\input{chapters/Chapter4}



\chapter{Implementation \& Testing}
\input{chapters/Chapter5}


\chapter{System Integration}
\input{chapters/Chapter6}


% Conclusions / Findings
\chapter{Conclusion}
\input{chapters/conclusions}

% Reflective analysis
% end of thesis body
% --------------------------
% Print out the references
% ============================================================
%  BIBLIOGRAPHY
% ============================================================
\begin{thebibliography}{99}
	
	\bibitem{rfc791}
	J. Postel, \textit{Internet Protocol}, IETF RFC 791, September 1981.
	
	\bibitem{rfc793}
	J. Postel, \textit{Transmission Control Protocol}, IETF RFC 793, September 1981.
	
	\bibitem{rfc768}
	J. Postel, \textit{User Datagram Protocol}, IETF RFC 768, August 1980.
	
	\bibitem{ieee8023}
	IEEE Std 802.3-2022, \textit{IEEE Standard for Ethernet}, IEEE, 2022.
	
	\bibitem{xilinxzynq}
	Xilinx Inc., \textit{Zynq-7000 SoC Technical Reference Manual}, UG585, v1.13, 2022.
	
	\bibitem{xilinxaxidma}
	Xilinx Inc., \textit{AXI DMA LogiCORE IP Product Guide}, PG021, v7.1, 2021.
	
	\bibitem{riscvspec}
	A.~Waterman and K.~Asanovi\'c, \textit{The RISC-V Instruction Set Manual, Volume I: Unprivileged ISA}, RISC-V Foundation, 2019.
	
	\bibitem{nmap}
	G. Lyon, \textit{Nmap Network Scanning}, Insecure.com LLC, 2009.
	
	\bibitem{snort}
	M. Roesch, ``Snort: Lightweight intrusion detection for networks,'' in \textit{Proc. USENIX LISA}, 1999.
	
	\bibitem{patterson_hennessy}
	D. A. Patterson and J. L. Hennessy, \textit{Computer Organization and Design RISC-V Edition: The Hardware/Software Interface}, 2nd ed. Morgan Kaufmann, 2021.
	
	\bibitem{harris_harris}
	S. Harris and D. Harris, \textit{Digital Design and Computer Architecture: RISC-V Edition}. Morgan Kaufmann, 2022.
	
	\bibitem{hennessy_patterson_advanced}
	J. L. Hennessy and D. A. Patterson, \textit{Computer Architecture: A Quantitative Approach}, 6th ed. Morgan Kaufmann, 2019.
	
	\bibitem{ashenden}
	P. J. Ashenden, \textit{The Designer's Guide to VHDL}, 3rd ed. Morgan Kaufmann, 2008.
	
	\bibitem{pedroni}
	V. A. Pedroni, \textit{Circuit Design and Simulation with VHDL}, 2nd ed. MIT Press, 2010.
	
	\bibitem{ug892}
	AMD Xilinx, \textit{Vivado Design Suite User Guide: Design Flows Overview}, UG892, 2023.
	
	\bibitem{ds180}
	AMD Xilinx, \textit{7 Series FPGAs Data Sheet: Overview}, DS180, 2023.
	
	\bibitem{xpack_gcc}
	xPack Project, \textit{xPack GNU RISC-V Embedded GCC}. [Online]. Available: \url{https://xpack.github.io/riscv-none-elf-gcc/}
	
	\bibitem{gnu_ld}
	Free Software Foundation, \textit{Using ld: The GNU Linker}. [Online]. Available: \url{https://sourceware.org/binutils/docs/ld/}
	
	\bibitem{gnu_as}
	Free Software Foundation, \textit{Using as: The GNU Assembler}. [Online]. Available: \url{https://sourceware.org/binutils/docs/as/}
	
	\bibitem{modelsim}
	Siemens EDA, \textit{ModelSim User's Manual}. Mentor Graphics, 2023.
	
\end{thebibliography}
%\include{chapters/appendices}

\end{document}
